FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-18 NewVersion/sections/Overview.tex
Overview

This case is the so-called \textit{unintentional method conflict}. It arises when two inherited methods happen to have
the same name and parameter types, but they are designed for different functionalities with different semantics.
Now one may quickly come up with a workaround, which is to manually merge them, by
creating a new \lstinline|draw| method in \lstinline|DrawableDeck| to
override the old ones. However, merging two methods with totally different functionalities does not make any sense.
This non-solution will hide the
old methods and break independent extensibility.

@@

\paragraph{\MIM's solution:} 
The essential problem is how to ensure that the correct method is
invoked. To solve this problem, \MIM{} uses a variant of method
dispatching that we call \textit{hierarchical dispatching}. In
hierarchical dispatching both the static and dynamic type information 
are used to select the right method implementation.
%Inspired by C++, we have formalized this feature into a minimal core calculus,
%studied and proved the type soundness. It is concise and can readily be imported to other systems. 
%And in the formalization, we propose \textit{hierarchical dispatch} in our model
%as the default method lookup algorithm. 
During runtime, the method call \lstinline|e.m()|
makes use of both the static type and the dynamic type of the receiver, so it seems to be a
combination of static and dynamic dispatch. Intuitively, the static type specifies one branch
to avoid ambiguity, and the dynamic type finds the most specific implementation on that branch. The formal algorithm will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.

@@

// main program
new DrawableSafeDeck().shuffleAndDraw() // SafeDeck.draw is called
\end{lstlisting}\vspace{3pt}
\bruno{Marco complined about this but this was not addressed either!!!}
During the runtime, when \lstinline|this.draw()| is called,
the dynamic type of the receiver is \lstinline|DrawableSafeDeck|, but the static type is actually
\lstinline|Deck|. Here it is an instance of the \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| method along path \lstinline|Deck|''. The meaning of ``along path
\lstinline|Deck|'' not only implies what \lstinline|Deck| inherits, but also the domain of its subtypes up to the
dynamic type \lstinline|DrawableSafeDeck|. Finally \lstinline|SafeDeck.draw| is correctly returned.

%Note that in C++, if we put the \lstinline[language=c++]|virtual| keyword in front of every \lstinline|draw| method, that means
%we enforce the dynamic dispatch, and it does the disambiguation correctly.
%Unfortunately, the C++ language, or rather the feature on triangle inheritance, has not been well formalized.
In our language model, there are two types of casts: \emph{implicit casts} and \emph{explicit upcasts}. The implicit casts appear
when passing arguments and returning expressions, they are also
upcasts. Hence all casts are safe. \bruno{This discussion seems
  inconplete. Either complete it or remove it.}

@@

\subsection{Problem 3: Overriding on Individual Branches}\label{subsec:partialoverrides}
Method overriding is common in Object-Oriented Programming. 
With diamond inheritance where conflicting methods are intended to have 
the same semantics, method overriding is not a problem. If conflicting
methods arise from multiple parents, we can override all those methods 
in a single unified (or merged) method in the subclass. Therefore
further overriding is simple, because there is only one method that
can be overridden.

>>>>

This case is an example of a so-called \textit{unintentional method conflict}. It arises when two inherited methods happen to have
the same name and parameter types, but they are designed for different functionalities with different semantics.
Now one may quickly come up with a workaround, which is to manually
merge the two methods, by
creating a new \lstinline|draw| method in \lstinline|DrawableDeck| to
override the old ones. However, merging two methods with totally different functionalities does not make any sense.
This non-solution will hide the
old methods and break independent extensibility.

@@

\paragraph{\MIM's solution:} 
The essential problem is how to ensure that the correct method is
invoked. To solve this problem, \MIM{} uses a variant of method
dispatching that we call \textit{hierarchical dispatching}. In
hierarchical dispatching both the static and dynamic type information 
are used to select the right method implementation.
%Inspired by C++, we have formalized this feature into a minimal core calculus,
%studied and proved the type soundness. It is concise and can readily be imported to other systems. 
%And in the formalization, we propose \textit{hierarchical dispatch} in our model
%as the default method lookup algorithm. 
During runtime, the method call \lstinline|e.m()|
makes use of both the static type and the dynamic type of the
receiver, so it is a
combination of static and dynamic dispatch. Intuitively, the static type specifies one branch
to avoid ambiguity, and the dynamic type finds the most specific implementation on that branch. The formal algorithm will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.

@@

// main program
new DrawableSafeDeck().shuffleAndDraw() // SafeDeck.draw is called
\end{lstlisting}\vspace{3pt}
\bruno{Marco complined about this but this was not addressed either!!!}
During the runtime, when \lstinline|this.draw()| is called,
the dynamic type of the receiver is \lstinline|DrawableSafeDeck|, but the static type is actually
\lstinline|Deck|. This is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| method along path \lstinline|Deck|''. The meaning of ``along path
\lstinline|Deck|'' not only implies what \lstinline|Deck| inherits, but also the domain of its subtypes up to the
dynamic type \lstinline|DrawableSafeDeck|. Finally \lstinline|SafeDeck.draw| is correctly returned.

%Note that in C++, if we put the \lstinline[language=c++]|virtual| keyword in front of every \lstinline|draw| method, that means
%we enforce the dynamic dispatch, and it does the disambiguation correctly.
%Unfortunately, the C++ language, or rather the feature on triangle inheritance, has not been well formalized.
%In our language model, there are two types of casts: \emph{implicit casts} and \emph{explicit upcasts}. The implicit casts appear
%when passing arguments and returning expressions, they are also
%upcasts. Hence all casts are safe. \bruno{This discussion seems
%  inconplete. Either complete it or remove it.}

@@

\subsection{Problem 3: Overriding on Individual Branches}\label{subsec:partialoverrides}
Method overriding is common in Object-Oriented Programming. 
With diamond inheritance, where conflicting methods are intended to have 
the same semantics, method overriding is not a problem. If conflicting
methods arise from multiple parents, we can override all those methods 
in a single unified (or merged) method in the subclass. Therefore
further overriding is simple, because there is only one method that
can be overridden.
